---
title: "@realDonaldTrump Tweets for Dow Jones Index Direction Prediction"
author: Dmitry Petukhov
output:
  github_document:
    toc: true
    toc_depth: 2
---


__Prepare dataset contains:__

* __@realDonaldTrump tweets,__

* __its sentiment scores from Microsoft Cognitive Service,__

* __Dow Jones NYSE index.__


Import dependencies and read config:
```{r set_env_options, include=FALSE}
options(max.print = 1e3, scipen = 999, width = 1e2)
options(stringsAsFactors = F)
```

```{r import_dependencies, include=FALSE}
suppressPackageStartupMessages({
  # data manupulation
  library(data.table)
  library(dplyr)
  library(tidyr)
  library(purrr)
  library(magrittr)

  # text processing
  library(stringr)
  library(tidytext)
  library(wordcloud)
  
  # other
  library(lubridate)
  library(Quandl)
})

```

```{r read_secrets, include=FALSE}
secrets <- config::get(file = "secrets.yml")
```

# Part I. Naive approach

## Get @realDonaldTrump Tweets

Download tweets:
```{r read_tweets, cache=TRUE}
donald_tweets_url <- sprintf("https://datainstinct.blob.core.windows.net/twitter/realdonaldtrump.csv?%s", secrets$azure_storage_key)


donald_tweets <- fread(donald_tweets_url, quote = "")

donald_tweets %<>% 
  filter(!is_retweet) %>% 
  mutate(
    id = as.character(id_str),
    
    # dates processing
    created_at = mdy_hms(created_at),
    created_date = as.Date(created_at),
    
    # text processing
    text = str_replace_all(text, '\n', " || "),
    text = str_replace_all(text, "&amp;", "and"),
    text = str_replace_all(text, '"', ""),
    text = str_replace_all(text, "http(s)://t.co/[A-Za-z\\d]+", "<url>"),
  ) %>% 
  select(-id_str)


donald_tweets %>% 
  select(created_at, text, retweet_count, favorite_count) %>% 
  arrange(desc(created_at)) %>% 
  as_tibble()
```

## Get Dow Jones NYSE index

Get [Dow Jones NYSE](https://www.quandl.com/data/BCB/7809-Dow-Jones-NYSE-index) data:

```{r get_dow_jones_index, cache=TRUE}
Quandl.api_key(secrets$quandl_api_key)


dji <- Quandl("BCB/7809")

dji %<>% 
  arrange(Date) %>% 
  mutate(diff = Value - lag(Value)) %>% 
  na.omit %>% 
  mutate(direction = if_else(diff < 0, 0L, 1L)) %>% 
  rename(date = Date)


dji %>% 
  arrange(desc(date)) %>% 
  as_tibble
```


## Join datasets

Aggregate tweets (daily):
```{r aggregate_twees}
donald_tweets_daily <- donald_tweets %>% 
  # calc daily stats
  group_by(created_date) %>% 
  summarise(
    n = n(),
    text = paste(text, collapse = " ||| ")
  ) %>% 
  ungroup() %>% 
  # join w/ Dow Jones Index
  inner_join(
    dji %>% select(date, direction), 
    by = c("created_date" = "date")
  ) %>% 
  # final touch
  mutate(direction = as.factor(direction))


donald_tweets_daily %>%
  arrange(desc(created_date)) %>% 
  as_tibble
```


## Save artifacts
```{r save_artifacts, include=FALSE}
write.table(donald_tweets_daily, 
            "cache/donald_tweets_daily.naive.csv",
            sep = ",", row.names = F, fileEncoding = "utf-8")
```

```{r gc, include=FALSE}
rm(donald_tweets_daily)
```



# Part II. NLP approach

## Tweets sentiment 

Get tweets sentiment score from Azure Cognitive Services [Text Analytics API](https://azure.microsoft.com/en-us/services/cognitive-services/text-analytics/):  
```{r get_sentiment, cache=TRUE}
source("cognitive_services_api.R")

donald_tweets_sentiments <- donald_tweets %>%
  filter(year(created_date) == 2019 & month(created_date) == 6) %>% # TODO: only for debug
  group_by(lubridate::year(created_date), lubridate::month(created_date)) %>% 
  group_map(~ get_sentiment(.x, secrets$cognitive_services_api_key, verbose = T))

donald_tweets_sentiments %<>% 
  bind_rows %>% 
  transmute(id, sentiment_score = score)


donald_tweets %>% 
  inner_join(donald_tweets_sentiments, by = "id") %>%
  select(text, sentiment_score) %>% 
  as_tibble
```



## Join datasets

Aggregate tweets (daily):

```{r donald_tweets_daily}
donald_tweets_daily <- donald_tweets %>% 
  # join w/ tweets sentiments
  inner_join(donald_tweets_sentiments, by = "id") %>% 
  # calÑ daily aggregates
  group_by(created_date) %>% 
  summarise(
    n = n(),
    
    sentiment_score_min = min(sentiment_score),
    sentiment_score_mean = mean(sentiment_score),
    sentiment_score_max = max(sentiment_score),
    
    retweet_count_total = sum(retweet_count),
    retweet_count_median = median(retweet_count),
    
    favorite_count_total = sum(favorite_count),
    favorite_count_median = median(retweet_count),
    
    text_total = paste(text, collapse = " ||| ")
  ) %>% 
  ungroup() %>% 
  # join w/ DJI
  inner_join(
    dji %>% select(date, direction), 
    by = c("created_date" = "date")
  ) %>% 
  # final touch 
  mutate(direction = as.factor(direction)) %>% 
  rename(text = text_total)


donald_tweets_daily %>% as_tibble
```


```{r save_to_cache, include=FALSE}
saveRDS(donald_tweets_daily, sprintf("cache/processed.%s.rds", Sys.Date()))
```


## Remove stopwords

```{r remove_stopwords}
tweet_words <- donald_tweets_daily %>%
  select(created_date, text) %>% 
  unnest_tokens(word, text, token = "regex", pattern = "([^A-Za-z\\d#@']|'(?![A-Za-z\\d#@]))") %>% # remove @ and #
  filter(
    !word %in% stop_words$word &
    str_detect(word, "[a-z]")
  )


tweet_words %>% 
  count(word) %>% 
  arrange(desc(n))
```


## Stats and word cloud
```{r word_clouds}
tweet_word_counts <- tweet_words %>% count(created_date, word)


# go to 4-th page
tweet_word_counts %>% 
  count(word) %>% 
  arrange(-n)

# go to 6-th page
tweet_word_counts %>% 
  filter(year(created_date) == 2019) %>% 
  count(word) %>% 
  arrange(-n)


total_words <- tweet_word_counts %>% 
  group_by(created_date) %>% 
  summarize(total = sum(n))

total_words %>% arrange(-total)


tweet_words %>%
  count(word) %>%
  with(wordcloud(word, n, max.words = 50, color = "gray"))
```


## Compute TD-IDF
```{r calc_td_idf}
tweet_word_counts <- tweet_word_counts %>% 
  left_join(total_words, by = "created_date") %>%
  bind_tf_idf(word, created_date, n)

tweet_word_counts %>% 
  group_by(created_date) %>% 
  filter(tf_idf == max(tf_idf)) %>% 
  arrange(desc(created_date))
```


## Calc Document-Term Matrix

```{r calc_dtm}
tweets_dtm <- cast_sparse_(tweet_word_counts, "created_date", "word", "tf_idf")

print(dim(tweets_dtm))
print(tweets_dtm[1:7, 1:7])
```


```{r view_result}
# TODO: remove 'direction'

donald_tweets_daily %<>% 
  select(-text) %>% 
  cbind(tweets_dtm %>% as.matrix %>% as_tibble)


donald_tweets_daily %>% as_tibble
```



```{r save_output, include=FALSE}
write.table(donald_tweets_daily, file = "cache/donald_tweets_daily.nlp.csv",
            sep = ",", fileEncoding = "utf-8",
            row.names = F, col.names = T, append = F, quote = T)
```

